---
title: "Untitled"
author: "Kelly Bodwin"
date: "1/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
library(tidyverse)
library(stringr)
library(readxl)
library(here)
```

```{r}
library(dbplyr)
library(odbc)
library(DBI)
library(RMySQL)
library(rsample)
library(glue)
```

`dbplyr` is the one that lets you use `dplyr` R code to sneakily run SQL queries.  

The rest of them are just "some packages I need for SQL" and I haven't bothered to understand what tasks each one covers, or even which of the functions I'm about to use come from which package.


```{r}
con <- dbConnect(RMySQL::MySQL(), 
                 user='kbodwin', 
                 password='vineyard', 
                 dbname='vineyard_diversity',
                 host='vineyard-project.cnmgssyjqpsv.us-west-1.rds.amazonaws.com',
                 port=3306)


```

This will show you what datasets are available in the database:

I changed this to used DBL's since the dplyr one was not working for me and I found a thread with hadley recommending this version. 

```{r}
dbListTables(con)
```

This will look like it makes a dataset in R, but in fact, it's just a pointer to a connection:

```{r}
df <- tbl(con, "Counts_by_Class")

df %>% arrange(desc(Class))
```

To get the dataset into R, you use `collect()`:

```{r}
counts_by_class <- df %>% collect()
```

The goal here is to "outsource" as much of the large-scale computation as possible to SQL, before collecting to R.

I think the best approach is to pre-establish the permutation rules, and then make SQL do the grouping and summarizing.

Something like:

```{r}
df %>%
  mutate(
    Experiment = case_when(
      treatment %in% c("NF", "OF", "SF") ~ "Fert",
      treatment %in% c("HERB", "NH") ~ "Herb",
      treatment %in% c("NCC", "HWD", "LWD") ~ "CC"
    )) %>%
  group_by(Experiment, season, year, Class) %>%
  # mutate(
  #   num = tally(),
  #   labels = sample(1:n, n)
  # ) %>%
  show_query()
```

(https://cran.r-project.org/web/packages/tidypredict/vignettes/sql.html)



```{r}
# Goals of get_perm_pvals
## The function iterate across all selected treatement columns 
## For the function to not repeat itself more then necessary, 
### for example if A is compared to B, B does not need to be comapred to A
## For the final output to be a dataframe containing all comparisons and p-values 
# General Notes
## data must include permutation 0 

get_perm_pvals <- function(data, class, ...){
  pvals <- data.frame() #initializes an empty dataframe
  col_vars <- quos(...)
  for(i in 1:length(col_vars)){    #iterates across list of treatments
    for(l in (i+1):length(col_vars)){   #iterates across list of treatments again
      if(l > length(col_vars)){
        return(pvals)
      }
      pvals_temp <- data %>%   
        mutate_if(is.numeric, ~replace_na(., 0)) %>%
        mutate(treat_diff = (!!!col_vars[i]) - (!!!col_vars[l]))
      pvals_temp <- pvals_temp %>% #creates difference column for each pair
        mutate(`x1-x2` = glue("{col_vars[i]}-{col_vars[l]}"))
      pvals_temp <- pvals_temp %>%                                        #creates column for each treatment pair
        group_by(season, {{class}}) %>%                       
        summarize(
          p_val = (sum(abs(treat_diff) >= abs(treat_diff[1]))-1)/1000,  # minus one because the real data equals itself
          treat_diff = treat_diff[1],
          `x1-x2` = `x1-x2`[1]) %>%
        arrange(p_val, treat_diff)
      pvals <- bind_rows(pvals_temp, pvals)
    }
  }
  return(pvals)
}

```

```{r}
add_Perm_ID <- function(perm_dat){
  new_perm <- NULL
  for(i in 1:nrow(perm_dat)){
   temp <- perm_dat$splits[[i]] %>% 
      analysis() %>% 
      mutate(Perm = perm_dat$id[i])
   new_perm <- rbind(new_perm, temp)
  }
  return(new_perm)
}
```

```{r}
class_df <- df %>% 
  filter(treatment %in% c("NF", "OF", "SF"),
         season == "Bloom") %>% 
  group_by(Class, treatment, season) %>%
  summarize(count = sum(Count)) %>% 
  collect()

class_df <- class_df %>% 
  pivot_wider(names_from = treatment, values_from = count) 
```

```{r}
class_perm <- class_df %>% 
  permutations(c("SF", "OF", "NF"), times = 1000, apparent = TRUE)

class_perm <- add_Perm_ID(class_perm)

class_perm <- class_perm[order(class_perm$Perm),]
```

```{r, warning=FALSE}
#debugonce(get_perm_pvals)
perm_pvals <- class_perm %>% 
  get_perm_pvals(class = Class, SF, OF, NF)
```

```{r}
perm_pvals %>% 
  filter(p_val <= 0.05) %>% 
  arrange(p_val)

perm_pvals$`x1-x2` %>% unique()
```
There are 22 significantly different counts by Class for the fertilization treatments in Bloom.  

One optimization that needs to be made is to eliminate self-comparison. Probably just need and if statement in the function. 


```{r}
df <- tbl(con, "Counts_by_Species")

species_df <- df %>% 
  filter(treatment %in% c("NF", "OF", "SF"),
         season == "Bloom") %>% 
  group_by(Species, treatment, season) %>%
  summarize(count = sum(Count)) %>% 
  collect()

species_df <- species_df %>% 
  pivot_wider(names_from = treatment, values_from = count) 
```

Need to switch to using data.tables, rbinding 1000 semi-large datasets takes way to much time. This piece of code is definitley going to be the bottle neck for the entire process. The code below took about took ~10 mins to run. 

```{r}
species_perm <- species_df %>% 
  permutations(c("SF", "OF", "NF"), times = 1000, apparent = TRUE)

tictoc::tic()
species_perm <- add_Perm_ID(species_perm)
tictoc::toc()

#To order the dataset from original data down to permutation 1000
species_perm <- species_perm[order(species_perm$Perm),]
```

The code below took ~8 mins to run. How to optimize it? 
```{r, warning=FALSE}
#debugonce(get_perm_pvals)
tictoc::tic()
perm_pvals <- species_perm %>% 
  get_perm_pvals(class = Species, SF, OF, NF)
tictoc::toc()
```

```{r}
perm_pvals %>% 
  filter(p_val <= 0.05) %>% 
  arrange(p_val)

perm_pvals$`x1-x2` %>% unique()
```

```{r}
df <- tbl(con, "Counts_by_Class")

counts_by_class <- df %>% collect()
```

```{r}
actinobacteria_df <- df %>% 
  filter(treatment %in% c("NF", "OF", "SF"),
         season == "Bloom",
         Class == "actinobacteria")  %>% 
  group_by(Class, treatment, season, sample) %>%
  summarise(Count = sum(Count)) %>% 
  collect()
```

```{r}
actinobac_perm <- actinobacteria_df %>% 
  permutations("treatment", times = 1000, apparent = TRUE)

actinobac_perm <- add_Perm_ID(actinobac_perm)

actinobac_perm <- actinobac_perm[order(actinobac_perm$Perm),]

actinobac_perm <- actinobac_perm %>% 
  group_by(Class, season, treatment, Perm) %>% 
  summarise(Count = mean(Count)) %>% 
  pivot_wider(names_from = treatment, values_from = Count) 
```

```{r, warning=FALSE}
#debugonce(get_perm_pvals)
perm_pvals <- actinobac_perm %>% 
  get_perm_pvals(class = Class, SF, OF, NF)
```




















